MSB1015_Assignment 2

### 2019-10-12 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved documentation of section 1 + added references
* The documentation of section 1, on the SPARQL query, was outdated and insufficient. This has been updated now. 
* The references for the WikiData and CDK databases were missing, these have been added.

In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData. I use the smiles to get descriptors from the CDK database. These descriptors contain information on the structural properties of the alkanes (see section 2 for more details). Finally, I train a Partial Least Squares (PLS) model to predict these properties from the chemical properties of the compounds and plot the results.

@@

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData. I get their name, boiling point, boiling point unit and smiles.

@@

@@

@@

**7. References**
ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.

>>>>

In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData (ref2). I use the smiles to get descriptors from the CDK database (ref3-6). These descriptors contain information on the structural properties of the alkanes (see section 2 for more details). Finally, I train a Partial Least Squares (PLS) model to predict these properties from the chemical properties of the compounds and plot the results.

@@

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to run a SPARQL query that obtains all simple alkines with known boiling points and smiles from WikiData.

@@

#The function query_wikidata runs the SPARQL query to obtain data from the Wikidata database. This returns a dataframe containing the alkane names, boiling points, units and smiles. It also contains the wikidata-links for compounds and units. There are 134 alkanes at the time of writing.

@@

```
Description of WikiData numbers:
* P31 = instance of
* P279 = subclass of
* Q41581 = alkane
* P2102 = boiling point
* P233 = canonical smiles

The SPARQL query can easily be changed to obtain different data. For example, one could get the smiles of alkynes instead of alkanes by changing Q41581 into Q159226. Look at https://www.wikidata.org/wiki/Wikidata:Main_Page to find the right numbers.

@@

**7. References**
ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.
ref2: https://www.wikidata.org/wiki/Wikidata:Main_Page (12-10-2019)
ref3: Willighagen et al. The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching. J. Cheminform. 2017; 9(3), doi:10.1186/s13321-017-0220-4
ref4: May and Steinbeck. Efficient ring perception for the Chemistry Development Kit. J. Cheminform. 2014, doi:10.1186/1758-2946-6-3
ref5: Steinbeck et al. Recent Developments of the Chemistry Development Kit (CDK) - An Open-Source Java Library for Chemo- and Bioinformatics. Curr. Pharm. Des. 2006; 12(17):2111-2120, doi:10.2174/138161206777585274
ref6: Steinbeck et al. The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics. J. Chem. Inf. Comput. Sci. 2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y
